# Data transformation

Originally, our team collected seven different cryptocurrency stocks (`ADA`, `BTC`, `DOGE`, `ETC`, `ETH`, `SOL`, `XRP`). One file for each stock contains the information of stocks, such as `Date`, `Open`, `High`, `Low`, `Close`, `Adj Close` and `Volume`. After investigating the missing data pattern, as described in the `Missing values` section, we decided to discard the `SOL` data in the analysis. In the following discussion, since we won't be using `Close` values in the analysis, we called `Adj Close` to be `Close` for convenience. There are several different dataframes we used in the later analysis. We will briefly introduce each of them and show the structure of each dataframe. 

## Close & Volume
The first dataframe is mainly used to analyze whether there is relationship among `Adj Close` and `Volume`of each stock. And we merged these three columns of each stock together in the same data frame by the date. The structure of the dataframe looks like:

```{r}
library(plyr)
library(lubridate)
library(tidyverse)
ADA <- data.frame(read.csv(file = './data/ADA-USD.csv')[, c("Date","Adj.Close", "Volume")])
BTC <- data.frame(read.csv(file = './data/BTC-USD.csv')[, c("Date","Adj.Close", "Volume")])
DOGE <- data.frame(read.csv(file = './data/DOGE-USD.csv')[, c("Date","Adj.Close", "Volume")])
ETC <- data.frame(read.csv(file = './data/ETC-USD.csv')[, c("Date","Adj.Close", "Volume")])
ETH <- data.frame(read.csv(file = './data/ETH-USD.csv')[, c("Date","Adj.Close", "Volume")])
XRP <- data.frame(read.csv(file = './data/XRP-USD.csv')[, c("Date","Adj.Close","Volume")])

dataName = c("ADA", "BTC", "DOGE", "ETC", "ETH", "XRP")
for (name in dataName) {
  lapply(get(name)[["Adj.Close"]],as.numeric)
}

names(ADA)[2] <- paste("ADA", "Adj.Close", sep="_")
names(ADA)[3] <- paste("ADA", "Volume", sep="_")
names(BTC)[2] <- paste("BTC", "Adj.Close", sep="_")
names(BTC)[3] <- paste("BTC", "Volume", sep="_")
names(DOGE)[2] <- paste("DOGE", "Adj.Close", sep="_")
names(DOGE)[3] <- paste("DOGE", "Volume", sep="_")
names(ETC)[2] <- paste("ETC", "Adj.Close", sep="_")
names(ETC)[3] <- paste("ETC", "Volume", sep="_")
names(ETH)[2] <- paste("ETH", "Adj.Close", sep="_")
names(ETH)[3] <- paste("ETH", "Volume", sep="_")
names(XRP)[2] <- paste("XRP", "Adj.Close", sep="_")
names(XRP)[3] <- paste("XRP", "Volume", sep="_")

cryptos <- join_all(list(ADA, BTC, DOGE, ETC, ETH, XRP), by='Date', type='full')
cryptos$Date = as.Date(cryptos$Date)
```

## Return & Standardized Return

In one of the analysis, the team tried to investigate in the relationship among the stocks in the perspective of their daily return. The daily return was calculated by: $return = (today's\;close - yesterday's\;close) / yeterday's\;close$. The team performed the analysis on both original return and the standardized return. The original return version looks like:
```{r}
close=cryptos[grep(".*_Adj.Close",colnames(cryptos))]
close_date = data.frame(Date=cryptos$Date,close)
close_scaled=data.frame(scale(close))
close_date_scaled=data.frame(Date=cryptos$Date,close_scaled)
return = data.frame(apply(close,2,diff,1,1))

i=0
for (name in colnames(return)){
  i=i+1
  colnames(return)[i]=paste(substr(name,1,3),"return",sep="_")
}
return_date = data.frame(Date=cryptos$Date[2:nrow(cryptos)],return)
str(return_date)
```

In the dataframe with standardized daily return, we standardized the daily return for each stock to perform a better analysis. The only difference compared to the previous dataframe is that for each column except the `Year` and `Month` column, we stored the standardized value. The dataframe is as shown below:
```{r}
return_scaled = data.frame(scale(return))
return_date_scaled = data.frame(Date=cryptos$Date[2:nrow(cryptos)],return_scaled)
str(return_date_scaled)
```

## Monthly Return and Standardized Monthly Return

The team tried to perform the analysis on monthly basis. In order to achieve the analysis, the set of dataframes with `Year`, `Month`, `Monthly Average Close` and the `Standardized Month Average Close` are used. The columns are as shown below:
```{r}
close_month = data.frame(Year=fct_inorder(factor(year(cryptos$Date))), Month=fct_inorder(factor(month(cryptos$Date, label=TRUE))),close) %>%
  dplyr::group_by(Year,Month)%>%
  dplyr::summarise(
    ADA_Adj.Close=mean(ADA_Adj.Close),
    BTC_Adj.Close=mean(BTC_Adj.Close),
    DOGE_Adj.Close=mean(DOGE_Adj.Close),
    ETC_Adj.Close=mean(ETC_Adj.Close),
    ETH_Adj.Close=mean(ETH_Adj.Close),
    XRP_Adj.Close=mean(XRP_Adj.Close)
    )%>%
  ungroup()

str(close_month)
```

In the dataframe with standardized monthly closes, we standardized the monthly average prices for each stock to perform a better analysis. The structure of the previous dataframe looks almost the same as the second one. The only difference is that for each column except the `Year` and `Month` column, we store the standardized value instead of the originally monthly price. The dataframe is as shown below:
```{r}
close_month_scaled = data.frame(Year=fct_inorder(factor(year(cryptos$Date))), Month=fct_inorder(factor(month(cryptos$Date, label=TRUE))),close_scaled) %>%
  dplyr::group_by(Year,Month)%>%
  dplyr::summarise(
    ADA_Adj.Close=mean(ADA_Adj.Close),
    BTC_Adj.Close=mean(BTC_Adj.Close),
    DOGE_Adj.Close=mean(DOGE_Adj.Close),
    ETC_Adj.Close=mean(ETC_Adj.Close),
    ETH_Adj.Close=mean(ETH_Adj.Close),
    XRP_Adj.Close=mean(XRP_Adj.Close)
    )%>%
  ungroup()

str(close_month_scaled)
```

## Close & Standardized Close

The last set of dataframes we used is quite simple. We include the date and the closing prices for each stock in one data frame. The structure looks like:

```{r}
ADA <- data.frame(read.csv(file = './data/ADA-USD.csv')[, c("Date","Adj.Close")])
BTC <- data.frame(read.csv(file = './data/BTC-USD.csv')[, c("Date","Adj.Close")])
DOGE <- data.frame(read.csv(file = './data/DOGE-USD.csv')[, c("Date","Adj.Close")])
ETC <- data.frame(read.csv(file = './data/ETC-USD.csv')[, c("Date","Adj.Close")])
ETH <- data.frame(read.csv(file = './data/ETH-USD.csv')[, c("Date","Adj.Close")])
XRP <- data.frame(read.csv(file = './data/XRP-USD.csv')[, c("Date","Adj.Close")])

dataName = c("ADA", "BTC", "DOGE", "ETC", "ETH", "XRP")
for (name in dataName) {
  lapply(get(name)[["Adj.Close"]],as.numeric)
}

names(ADA)[2] <- paste("ADA", "Adj.Close", sep="_")
names(BTC)[2] <- paste("BTC", "Adj.Close", sep="_")
names(DOGE)[2] <- paste("DOGE", "Adj.Close", sep="_")
names(ETC)[2] <- paste("ETC", "Adj.Close", sep="_")
names(ETH)[2] <- paste("ETH", "Adj.Close", sep="_")
names(XRP)[2] <- paste("XRP", "Adj.Close", sep="_")

cryptos <- join_all(list(ADA, BTC, DOGE, ETC, ETH, XRP), by='Date', type='full')
cryptos$Date = as.Date(cryptos$Date)
```

```{r}
str(cryptos)
```

In the dataframe with standardized close, instead of containing the original stock prices, we standardized the prices for each stock to perform a better analysis. The structure of the last dataframe looks almost the same as the second one. The only difference is that for each column except the `Date` column, we store the standardized value instead of the originally price. The dataframe is as shown below:

```{r}
cryptos <- join_all(list(ADA, BTC, DOGE, ETC, ETH, XRP), by='Date', type='full')
cryptos$Date = as.Date(cryptos$Date)
close=cryptos[grep(".*_Adj.Close",colnames(cryptos))]
close_date = data.frame(Date=cryptos$Date,close)
close_scaled=data.frame(scale(close))
close_date_scaled=data.frame(Date=cryptos$Date,close_scaled)
```

```{r}
str(close_date_scaled)
```
